\bf Classification and Regression on Random Dot Product Graphs

### 2024-02-22 draft.Rmd
forgot to add a few things

abstract: |
  The random dot product graph (RDPG) has become a powerful modeling tool in uncovering latent structures within graphs. In particular, it has been shown that the RDPG describes a wide range of popular random graph models with rigid latent structures. More recently, joint modeling of mutliple random graphs that share common properties or structures across graphs have been introduced, such as the multilayer RDPG, multiple RPDG, multilayer stochastic block model, etc. In this work, we use these joint random graph models in the context of statistical learning, such as classification and regression, by introducing the multiple latent structure model, in which the graphs share a common latent structure with different parameters that correspond to different response variables. Then we propose various estimation techniques involving manifold learning to estimate these parameters and in turn predict the responses, with theorems guaranteeing convergence of the predictions. Simulations, as well as applications on brain connectivity networks, verify the performance of our methods. 
---

@@

import::from(magrittr, `%>%`, `%<>%`)
import::from(foreach, foreach, `%do%`, `%dopar%`)
library(ggplot2)
library(ggraph)
source('~/dev/pabm-grdpg/functions.R')
source('~/dev/manifold-block-models/functions.R')

@@

>>>>

abstract: |
  The random dot product graph (RDPG) has become a powerful modeling tool in uncovering latent structures within graphs. In particular, it has been shown that the RDPG describes a wide range of popular random graph models with rigid latent structures. More recently, joint modeling of mutliple random graphs that share common properties or structures across graphs have been introduced, such as the multilayer RDPG, multiple RPDG, and multilayer stochastic block model. In this work, we use these joint random graph models in the context of statistical learning, such as classification and regression, by introducing the multiple latent structure model, in which the graphs share a common latent structure with different parameters that correspond to different response variables. Then we propose various estimation techniques involving manifold learning to estimate these parameters and in turn predict the responses, with theorems guaranteeing convergence of the predictions. Simulations, as well as applications on brain connectivity networks, verify the performance of our methods. 
---

@@

import::from(magrittr, `%>%`, `%<>%`)
import::from(foreach, foreach, `%do%`, `%dopar%`)
library(ggplot2)
library(ggraph)
library(mclust)
source('~/dev/pabm-grdpg/functions.R')
source('~/dev/manifold-block-models/functions.R')

@@

Next, we considered each hemisphere separately and fit a Beta distribution to the embedding of each half-network. 
The scatterplot of the fitted $\alpha$ and $\beta$ parameters suggests that there are two clusters of brain networks (figure \ref{fig:hcp-clustering}). 
Greater $\alpha$ values correspond to more embedding vectors farther away from the origin, corresponding to more nodes of higher degree, which in turn correspond to more brain regions with a greater number of connections. 
Conversely, greater $\beta$ values correspond to more embedding vectors closer to the origin, which in turn correspond to more brain regions with fewer connections. 
This may suggest that there are underlying groups, one of which tends to have higher connectivity in the left hemisphere, and the other having higher connectivity in the right hemisphere.

```{r hcp-clustering, cache = TRUE, fig.cap = 'Scatterplot of fitted parameters of the Beta distribution. The left image are the parameters fitted on the embedding of the left hemisphere, and the right image are the parameters fitted on the embedding of the right hemisphere.', fig.width = 8, fig.height = 4}
beta.df <- ase.stats.df[c('a.l', 'a.r', 'b.l', 'b.r')]

zhat.l <- mclust::Mclust(beta.df[, c('a.l', 'b.l')], G = 2)$classification
plot.l <- ggplot(beta.df) + 
  geom_point(aes(x = a.l, y = b.l, colour = factor(zhat.l))) + 
  coord_fixed() + 
  theme_bw() + 
  labs(x = expression(hat(alpha)[l]), y = expression(hat(beta)[l]), colour = 'cluster') + 
  scale_colour_brewer(palette = 'Set1')

zhat.r <- mclust::Mclust(beta.df[, c('a.r', 'b.r')], G = 2)$classification
plot.r <- ggplot(beta.df) + 
  geom_point(aes(x = a.r, y = b.r, colour = factor(zhat.r))) + 
  coord_fixed() +
  theme_bw() + 
  labs(x = expression(hat(alpha)[r]), y = expression(hat(beta)[r]), colour = 'cluster') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(plot.l, plot.r, nrow = 1)
```

### 2024-02-22 draft.tex
forgot to add a few things

\bigskip
\begin{abstract}
The random dot product graph (RDPG) has become a powerful modeling tool
in uncovering latent structures within graphs. In particular, it has
been shown that the RDPG describes a wide range of popular random graph
models with rigid latent structures. More recently, joint modeling of
mutliple random graphs that share common properties or structures across
graphs have been introduced, such as the multilayer RDPG, multiple RPDG,
multilayer stochastic block model, etc. In this work, we use these joint
random graph models in the context of statistical learning, such as
classification and regression, by introducing the multiple latent
structure model, in which the graphs share a common latent structure
with different parameters that correspond to different response
variables. Then we propose various estimation techniques involving
manifold learning to estimate these parameters and in turn predict the
responses, with theorems guaranteeing convergence of the predictions.
Simulations, as well as applications on brain connectivity networks,
verify the performance of our methods.
\end{abstract}

@@

>>>>

\bigskip
\begin{abstract}
The random dot product graph (RDPG) has become a powerful modeling tool
in uncovering latent structures within graphs. In particular, it has
been shown that the RDPG describes a wide range of popular random graph
models with rigid latent structures. More recently, joint modeling of
mutliple random graphs that share common properties or structures across
graphs have been introduced, such as the multilayer RDPG, multiple RPDG,
and multilayer stochastic block model. In this work, we use these joint
random graph models in the context of statistical learning, such as
classification and regression, by introducing the multiple latent
structure model, in which the graphs share a common latent structure
with different parameters that correspond to different response
variables. Then we propose various estimation techniques involving
manifold learning to estimate these parameters and in turn predict the
responses, with theorems guaranteeing convergence of the predictions.
Simulations, as well as applications on brain connectivity networks,
verify the performance of our methods.
\end{abstract}

@@

Next, we considered each hemisphere separately and fit a Beta
distribution to the embedding of each half-network. The scatterplot of
the fitted \(\alpha\) and \(\beta\) parameters suggests that there are
two clusters of brain networks (figure \ref{fig:hcp-clustering}).
Greater \(\alpha\) values correspond to more embedding vectors farther
away from the origin, corresponding to more nodes of higher degree,
which in turn correspond to more brain regions with a greater number of
connections. Conversely, greater \(\beta\) values correspond to more
embedding vectors closer to the origin, which in turn correspond to more
brain regions with fewer connections. This may suggest that there are
underlying groups, one of which tends to have higher connectivity in the
left hemisphere, and the other having higher connectivity in the right
hemisphere.

{\centering \includegraphics{draft_files/figure-latex/hcp-clustering-1}

\caption{Scatterplot of fitted parameters of the Beta distribution. The left image are the parameters fitted on the embedding of the left hemisphere, and the right image are the parameters fitted on the embedding of the right hemisphere.}\label{fig:hcp-clustering}
\end{figure}
